Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2019-12-31 analysis/paper/paper.Rmd
add some references

- the definition of craft specialization

@@

@@

social changes -  culture contact framework 
- the cause of emergence of craft specialization, elite control, rice intensification, political, economic 
- here I have the third cultural resistance, identity take a look at social boundary
- standardization as indicator, metric and technological
- GM is more sensitive for subtle changes, and the outline
- innovative: CV test 
- hypothesis, goals

@@

Several variables have been proposed to measure ceramic standardization, including metric, compositional, and technological attributes [@Rice1991; @Blackman1993; @Costin1991; @Hirshman2010; @Tite1999]. However, traditional typological and linear measurements are limited because they can be insensitive to subtle variations resulting from changes in craft specialization. Taking a new approach to the measurement of craft specialization, we studied shape and standardization of locally made pottery to identify changes in pottery production at Kiwulan (1400-1900 AD), a large multi-component archaeological site in northeastern Taiwan. We apply geometric morphometric methods to study artifact shapes to investigate if there are any changes resulting from foreign contact of European and Chinese that might indicate social changes in the indigenous society.

>>>>

@@

- the cause of emergence of craft specialization, elite control, rice intensification, political, economic

@@

- social changes -  culture contact framework 
- here I have the third cultural resistance, identity take a look at social boundary
- standardization as indicator, metric and technological
- GM is more sensitive for subtle changes, and the outline
- innovative: CV test 
- hypothesis, goals

@@

Several variables have been proposed to measure ceramic standardization, including metric, compositional, and technological attributes [@Rice1991; @Blackman1993; @Costin1991; @Hirshman2010; @Tite1999; @Eerkens2001]. However, traditional typological and linear measurements are limited because they can be insensitive to subtle variations resulting from changes in craft specialization. Taking a new approach to the measurement of craft specialization, we studied shape and standardization of locally made pottery to identify changes in pottery production at Kiwulan (1400-1900 AD), a large multi-component archaeological site in northeastern Taiwan. We apply geometric morphometric methods to study artifact shapes to investigate if there are any changes resulting from foreign contact of European and Chinese that might indicate social changes in the indigenous society.
